# Exploratory and descriptive analysis with R 


## Working example - record sales data


## Let's import the data {.smaller}

```{r}
Album_Sales <- read_csv("Datasets/album_sales.csv")

```

## Let's look at the data {.smaller}

```{r}

head(Album_Sales)
```
## Let's make sure our data types are correct #1 {.smaller}

- This variable is currently stored as charcters, not as a factor / category variable
```{r}

str(Album_Sales$Genre) 

```

## Let's make sure our data types are correct #2 {.smaller}

- We can  save it as a factor

```{r}
Album_Sales$Genre <- as.factor(Album_Sales$Genre)

str(Album_Sales$Genre) 
```

# Summarising data: Central tendency {.smaller}

## Measures of central tendency

The main measures of central tendency are:
- Mean
- Median
- Mode

## Mean {.smaller}

"What is the mean of album sales?"
  
```{r}
mean(Album_Sales$Sales)

```

## Trimmed mean {.smaller}

- The trimmed mean is used to reduce the influence of outliers on the summary

```{r}
mean(Album_Sales$Sales, trim = 0.05)

```

## Median {.smaller}

"What is the median amount of Airplay?"

```{r}

median(Album_Sales$Airplay)

```
## Mode {.smaller}

"What is the most common attractiveness rating of bands?"

- The easiest way to get the mode in R is to generate a frequency table

```{r}

table(Album_Sales$Attract)
```
- We can then look for the most frequently occuring response

# Measures of dispresion or variance {.smaller}

## Range  {.smaller}

The range is the difference between the lowest and highest values

- You can calculate it using these values

```{r}

max(Album_Sales$Airplay) - min(Album_Sales$Airplay)

```
 - Or you can use the range command to get the min and max values in one go

```{r}

range(Album_Sales$Airplay)

```

## Interquartile range {.smaller}

- We know that the median is the "middle" of the data = 50th percentile
- The interquatile range is the difference between the values at the 25th and 75th percentiles
```{r}
quantile( x = Album_Sales$Airplay, probs = c(.25,.75) )


```

- Interquartile range = 36 - 19.75 = 16.25

## Sum of squares {.smaller}

- The difference between each value and the mean value, squared, and then summed together

```{r}
sum( (Album_Sales$Adverts - mean(Album_Sales$Adverts))^2 )
```


## Variance {.smaller}

- Variance: Sum of sqaures divided by n-1

```{r}

# variance calculation
varianceAdverts <- sum( (Album_Sales$Adverts - mean(Album_Sales$Adverts))^2 ) / 199

```


## Standard deviation {.smaller}

- Standard deviation is square root of the variance

```{r}
# sd calculation


sqrt(varianceAdverts)


```
- Can be calculated using the sd() command

```{r}
sd(Album_Sales$Adverts)
```

## The *psych* package includes a lot of useful descriptive stats {.smaller}

```{r}
library("psych")
```

# Skewness and Kurtosis {.smaller}

## Assessing skewness of distribution #1 {.smaller}
- It is possible to use graphs to view the distribution
- We will focus on graphic presentation of data next week

```{r}
hist(Album_Sales$Sales)

```

## Assessing skewness of distribution #2 {.smaller}

- We can check raw skewness value using the *skew()* command in the **psych** package

```{r}
skew(Album_Sales$Sales)

```

## Kurtosis {.smaller}

| informal term |technical name|   kurtosis value |
|--:|--:|--:|
|"too flat"|platykurtic|   negative|
|  "just pointy enough"|  mesokurtic|  zero|
| "too pointy" |leptokurtic |positive |

```{r}
kurtosi(Album_Sales$Sales)
```

## Assessing normality of distribution {.smaller}
 
- We can use the shapiro-wilk test of normality
- This is part of "base" r (no package needed)

```{r}
shapiro.test(Album_Sales$Sales)
```

# Getting and overall summary {.smaller}

## summary() - in "base R" {.smaller}

```{r}
summary(Album_Sales)
```


## describe() - in the "psych" package #1 {.smaller}
```{r}
describe(Album_Sales)
```

## describe() - in the "psych" package #2 {.smaller}

- We can describe by factor variables 
```{r}
describeBy(Album_Sales, group = Album_Sales$Genre)
```


# Basic statistical tests (more detail in later sections)

## Corrleation {.smaller}

"Is there a relationship between advert spend and sales?"

- We would use an correlational analysis to answer this question

```{r}
plot(Album_Sales$Sales,Album_Sales$Adverts)

```



"Is there a relationship between advert spend and sales?"

- We would use an correlational analysis to answer this question

```{r}
cor.test(Album_Sales$Sales,Album_Sales$Adverts)


```

## Tests of difference - t-test {.smaller}

"Is there a significant difference in sales between the Country and Hip-hop musical genres?"

- We would use a t-test to answer this question

```{r}
myTtestData <- Album_Sales %>% filter(Genre == c("Country", "HipHop"))

t.test(myTtestData$Sales ~ myTtestData$Genre)

```
## Tests of difference - ANOVA {.smaller}

"Is there a significant difference in sales between all musical genres?"

- We would use an ANOVA to answer this question

```{r}
myAnova <- aov(Album_Sales$Sales ~ Album_Sales$Genre)
summary(myAnova)

```



