---
title: "Probability and Hypothesis Testing"
author: "Christopher J Wilson"
date: "13/10/2019"
output: ioslides_presentation
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(ggplot2)
library(dplyr)
library(Hmisc)
se <- function(x) sqrt(var(x)/length(x))
ci <- function(m,n) { qt(.975, df=n-1)*se(m) }
library(shiny)
```
# The role of probability in testing a hypothesis

## Let's imagine that 10 aliens invaded earth
- Most of them are quite tall
- Scientists hypothesise that the aliens might be taller than humans

## We know a lot about the height of the human population

- We know that the mean height of the human **population** is 
- We know that the standard deviation of the human **population** is
- We know that heights of the human **population** are normally distributed

# Because we know this at a population level, we can make exact statements about human height

## Distribution of heights in the human population
- This is called a probability distribution

# We only have a **sample** of alien heights

## We know a lot about the **sample**

- The mean of the alien **sample** heights is
- The standard deviation of alien **sample** heights is

# But how can we use this information to figure out what the alien **population** looks like?

## A t-distribution uses sample mean to estimate the population mean {.smaller}

- The t-distribution is a probability distribution
- Based on our sample mean and standard deviation, we can estimate the population mean

```{r}

##Two tailed

mean = 0
sd = 1
mydf <- 9
x=seq(-3,3,length=14)
y <- dt(x,df=mydf)



tdistPlot <- plot(x,y, type="l", col="red", lwd=2, main="t-distribution when sample size = 10",  xlab="Standard deviations")

alpha <- qt(0.025,df=mydf)
lb = sqrt(alpha*alpha)
ub = 3
 
i <- x >= lb & x <= ub
polygon(c(lb,x[i],ub), c(0,y[i],0), col="gray74") 
area <- pt(ub, df=mydf) - pt(lb, df=mydf)
lb = -3
ub = alpha
i <- x >= lb & x <= ub
polygon(c(lb,x[i],ub), c(0,y[i],0), col="gray74") 
area <- pt(ub, df=mydf) - pt(lb, df=mydf)


```

## A t-distribution uses sample mean to estimate the population mean
- The estimate changes based on:
  - Sample size
  - Standard deviation of the sample
  
- The estimate is more accurcate:
  - As the sample size gets bigger
  - As the standard deviation of the sample is lower

## Using the alien sample data to estimate alien population mean {.samller}

- Sample size =
- Mean  = 
- SD = 

```{r echo=FALSE}
##Two tailed

mean = 0
sd = 1
mydf <- 9
x=seq(-3,3,length=14)
y <- dt(x,df=mydf)



tdistPlot <- plot(x,y, type="l", col="red", lwd=2, main="t-distribution when sample size = 10",  xlab="Standard deviations")

alpha <- qt(0.025,df=mydf)
lb = sqrt(alpha*alpha)
ub = 3
 
i <- x >= lb & x <= ub
polygon(c(lb,x[i],ub), c(0,y[i],0), col="gray74") 
area <- pt(ub, df=mydf) - pt(lb, df=mydf)
lb = -3
ub = alpha
i <- x >= lb & x <= ub
polygon(c(lb,x[i],ub), c(0,y[i],0), col="gray74") 
area <- pt(ub, df=mydf) - pt(lb, df=mydf)


```


## Simulating sample data and confidence intervals 
```{r, echo = FALSE}

shinyApp(
# Define UI for application that draws a histogram
ui <- fluidPage(
   
   # Application title
   titlePanel("The effect of Sample Size, Mean and SD on the significance of the difference between groups"),
   
   # Sidebar with a slider input for number of bins 
   sidebarLayout(
      sidebarPanel(
        
        sliderInput("number", label = "Number of participants per group:",
                    min = 10, max = 1000, value = 50, step = 10),
        sliderInput("g1_mean", label = "Mean of group 1:",
                    min = 1, max = 100, value = 50, step = 1),
        sliderInput("g1_sd", label = "Standard deviation of group 1:",
                    min = 1, max = 100, value = 5, step = 1), 
        sliderInput("g2_mean", label = "Mean of group 2:",
                    min = 1, max = 100, value = 70, step = 1),
        
        sliderInput("g2_sd", label = "Standard deviation of group 2:",
                    min = 1, max = 100, value = 10, step = 1)
        
      ),
      
      # Show a plot of the generated distribution
      mainPanel(
         plotOutput("myPlot"),
         verbatimTextOutput("myt")
      )
   )
),


server <- function(input, output) {
   
 myData <- reactive({
   
   n <- input$number
   group1 <-  rep(1, n)
   group1Time <- rnorm(n, input$g1_mean, input$g1_sd)
   group2 <- rep(2,n)
   group2Time <- rnorm(n, input$g2_mean, input$g2_sd)
   group <- c(group1,group2)
   time <- c(group1Time, group2Time)
   hdata <- as.data.frame(cbind(group, time))
   hdata
 })
   
  output$myPlot <- renderPlot({
    
    myxlims <- c(min(myData()$time), max(myData()$time))
    
    p1 <- ggplot(myData(), aes(x=time, fill=as.factor(group))) + geom_histogram()  + theme(legend.position="none") + xlim(myxlims) + ggtitle("Histogram showing Group 1 and Group 2") + xlab("Time") + ylab("Frequency")
    
    p2 <- ggplot(myData(), aes(as.factor(group), time, colour=as.factor(group))) + stat_summary(fun.y = "mean", geom = "point") + stat_summary(geom="errorbar", fun.data=mean_cl_normal, width=0.1, conf.int=0.95) + coord_flip(ylim = c(myxlims)) + ggtitle("Confidence intervals of the mean") + xlab("Group") + ylab("Time") + theme(legend.position="none")
    
    
    
    gridExtra::grid.arrange(p1,p2, ncol=1)
    
  })
  
  output$myt <- renderPrint ({
    
    t.test(myData()$time~myData()$group)
    
  })
  
 
},

)

```