  ---
title: "Summarising Data"
author: "Christopher Wilson"
date: "7/25/2018"
output: ioslides_presentation
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(kableExtra)
library(sn)
```

# Summarising data: Frequencies

## One way to examine data is to look at the frequency distribution

- Frequencies = counting the frequency of a particular response.
- For example: We ask 100 people whether they like marmite.

```{r frequencies1}

N <- matrix(c(30,60,10),ncol=3, nrow = 1 )
kable(N, col.names = c("Yes","No","Not Sure"))


```

## Let's imagine we ask 100 people the question “How much time do you spend online each week?”

```{r online table}

Participant <- 1:250
Hours_Online <- rnorm(250, 8, 2.5)
kable(cbind(Participant, Hours_Online), "html", col.names = c("Participant", "Hours"))


```

## We can calculate the frequency of each response on a histogram

```{r online histogram}
library(ggplot2)
library(tidyr)

cbind.data.frame(Participant,Hours_Online) %>% ggplot(aes(x=Hours_Online)) + geom_histogram(bins=20, fill = "blue", colour="blue") # + geom_vline(aes(xintercept=mean(Hours_Online)),  color="red", linetype="dashed", size=1)
```

## The shape of disributions

``` {r }
 hist( rbeta(250, 5, 5), xlab = "Frequency", ylab = "Time spent online", main = "Normal distribution" )


```


## The shape of disributions

``` {r }

 hist( rbeta(250, 5, 2), xlab = "Frequency", ylab = "Time spent online", main = "Left-skewed distribution" )


```

## The shape of disributions

``` {r }
   hist( rbeta(250, 2, 5), xlab = "Frequency", ylab = "Time spent online", main = "Right-skewed distribution" )

```




## Interactive Plot

```{r eruptions}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```
